---
title: "liuKoladaMini: Minimal Kolada API Client"
author: "Mohan Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{liuKoladaMini: Minimal Kolada API Client}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(liuKoladaMini)
```

## Introduction

`liuKoladaMini` is a minimal client for the [Kolada API](https://www.kolada.se/),  
focusing on municipalities (kommuner) and indicators (KPI).  
It provides simple functions to explore and analyze Swedish municipality data.

Main functions:
- `kld_municipalities()` – list municipalities  
- `kld_indicators()` – list/search KPIs (indicators)  
- `kld_values()` – fetch values for a KPI in a municipality  
- `kld_plot()` – quick plotting of KPI values  

---

## Example Usage

### List municipalities
```{r}
head(kld_municipalities())
```

### Search indicators (e.g. "arbetslöshet")
```{r}
head(kld_indicators("arbetslöshet"))
```

### Fetch values (Linköping, KPI N02282: Arbetslöshet utrikes födda)
```{r}
df <- kld_values("0580", "N02282", 2019:2024)
df
```

### Plot values
```{r}
p <- kld_plot(df)
p
```

---

## Conclusion

This vignette demonstrates how to:
1. Find municipalities and indicators
2. Fetch KPI values
3. Plot trends quickly  

The package is designed as a lightweight educational example for Lab05.
